---
title: "Tire Choice (Conjoint) Analysis"
author: "Wedam Nyaaba"
date: "March 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Required Packages**

- ggplot2: For visualization of data
- conjoint: For market conjoint analysis

 &nbsp;
 &nbsp;
 
 \newpage
 
```{r}
# Clean current environment
rm(list = ls())

# Load conjoint package
require(conjoint)
```
Begin Analysis

# 1. Read in TireStudy Data Set for Preprocessing

```{r}
# Load Data
df<- read.csv("TireStudy.csv")

# Create Profiles (a combination of tire features) dataframe
 
profiles <- df[,2:5]
save(profiles, file="profiles.RData")
# Create customer rankings dataframe

custRanking <- t(df[,6:10])
save(custRanking, file="Rankings.RData")

colnames(custRanking) = c("profl-1","profl-2","profl-3","profl-4","profl-5","profl-6","profl-7",
                         "profl-8","profl-9","profl-10","profl-11","profl-12","profl-13","profl-14"
                         ,"profl-15","profl-16","profl-17","profl-18")
rownames(custRanking) = c("cust-1","cust-2","cust-3","cust-4","cust-5")

```
# 2. View Data Structure
```{r}
# Show profiles
profiles

str(profiles)
```

```{r}
# show custranking frame
head(custRanking)

str(custRanking)
```
# 3. Exploratory Analysis of Data

This is arbitrarily carried out for only customer 1  to observe how they rated tires based on their features: brand, price, life, and hazard. 

Here, it is good to add their respective rankings to the profiles dataframe.

```{r}
# Rename the profiles
combinedDF = profiles

# Add customer 1 and 2 rankings as new columns
combinedDF$cust1_Ranking <- as.vector(t(custRanking[1,]))
combinedDF$cust3_Ranking <- as.vector(t(custRanking[3,]))

str(combinedDF)
```
```{r}
# Load ggplot2 package
require(ggplot2)

```

```{r}
# Show boxplot of customer 1 ranking by Tire Brand 
ggplot(combinedDF, aes(x=Brand, y=cust1_Ranking, fill=Brand)) + geom_boxplot() + theme_classic() + labs(title = "Customer-1 Ranking by Tire Brand", x = "Brand", y = "Ranking") + theme(text= element_text(size=15)) + theme(plot.title = element_text(hjust=0.5))

```

```{r}
# Show boxplot of customer 1 ranking by Tire Price 
ggplot(combinedDF, aes(x=Price, y=cust1_Ranking, fill=Price)) + geom_boxplot() + theme_classic() + labs(title = "Customer-1 Ranking by Tire Price", x = "Price", y = "Ranking") + theme(text= element_text(size=15)) + theme(plot.title = element_text(hjust=0.5))
```
```{r}
# Show boxplot of customer 1 ranking by Tire Life 
ggplot(combinedDF, aes(x=Life, y=cust1_Ranking, fill=Life)) + geom_boxplot() + theme_classic() + labs(title = "Customer-1 Ranking by Tire Life", x = "Life (in miles)", y = "Ranking") + theme(text= element_text(size=15)) + theme(plot.title = element_text(hjust=0.5))
```
```{r}
# Show boxplot of customer 1 ranking by Tire Life 
ggplot(combinedDF, aes(x=Hazard, y=cust1_Ranking, fill=Hazard)) + geom_boxplot() + theme_classic() + labs(title = "Customer-1 Ranking by Tire Hazard", x = "Hazard", y = "Ranking") + theme(text= element_text(size=15)) + theme(plot.title = element_text(hjust=0.5))
```

# 4. Conjoint Analysis of Tire Rankings 

## 4.1 Customer 1 Conjoint Analysis

### 4.1.1 Obtain Attribute Importance

```{r}
# Create a level object that to hold categories of the features

level <- c(levels(combinedDF$Brand),levels(combinedDF$Price),levels(combinedDF$Life)
           ,levels(combinedDF$Hazard))
level
```

```{r}
# Calculate the % importance of each attribute (or feature)
FeatImp <- caImportance(y = custRanking[1,], x = profiles)
names(FeatImp) = names(profiles)
FeatImp
```
Visualize Feature Importance in a Bar plot
```{r}
barplot(FeatImp,col = "blue", width = 0.1, ylim = c(0, 60), xlab = "Attribute", ylab = "% Importance", main = "Feature/Attribute Importance")
```
### 4.1.2 Obtain part-worths/utilities (how much each product attribute contributes to the overall customer preference)

```{r}
partWorths <- caUtilities(y = combinedDF$cust1_Ranking, x = combinedDF[1:4], z = level)

```
```{r}
names(partWorths) = c("Intercept", level)

# Show part worths rounding to the nearest hundredth
formatC(partWorths[-1], format = 'f', digits = 2)
```
Visualize part-worths as follows.

```{r, fig.width=9, fig.height=7}
# Ceate a vector to assign the same color to categories of the same attribute
cols <- c(1,1,1,2,2,2,3,3,3,4,4)

# Horizontal barplot

plot <- barplot(partWorths[-1],las = 1, horiz = TRUE, xlim = c(-5, 7),axisnames = FALSE,
        col = cols, main = "Part-Worths (Utilities)")

text(x = 5, y= plot, cex = 1.0,
     labels= paste(level,
                   "(",
                   formatC(partWorths[-1], format = 'f', digits = 2),
                   ')'))
# Add legend

legend(-2.9, 13,toupper(names(profiles)), pch =0, cex = 1.0, fill = 1:length(FeatImp))


```

## 4.2 Customer-3 Conjoint Analysis

### 4.2.1 Obtain Attribute Importance


```{r}
# Calculate the % importance of each attribute (or feature)
FeatImp_3 <- caImportance(y = custRanking[3,], x = profiles)
names(FeatImp_3) = names(profiles)
FeatImp_3
```
Visualize Feature Importance in a Bar plot
```{r}
barplot(FeatImp_3,col = "blue", width = 0.1, ylim = c(0, 70), xlab = "Attribute", ylab = "% Importance", main = "Feature/Attribute Importance")
```
### 4.1.2 Obtain part-worths/utilities (how much each product attribute contributes to the overall customer preference)

```{r}
partWorths_3 <- caUtilities(y = combinedDF$cust3_Ranking, x = combinedDF[1:4], z = level)

```
```{r}
names(partWorths_3) = c("Intercept", level)

# Show part worths rounding to the nearest hundredth
formatC(partWorths_3[-1], format = 'f', digits = 2)
```
Visualize part-worths as follows.
```{r, fig.width=9, fig.height=7}
# Ceate a vector to assign the same color to categories of the same attribute
cols <- c(1,1,1,2,2,2,3,3,3,4,4)

# Horizontal barplot

plot <- barplot(partWorths_3[-1],las = 1, horiz = TRUE, xlim = c(-3, 5),axisnames = FALSE,
        col = cols, main = "Part-Worths (Utilities)")

text(x =4, y= plot, cex = 1.0,
     labels= paste(level,
                   "(",
                   formatC(partWorths_3[-1], format = 'f', digits = 2),
                   ')'))
# Add legend

legend(-2.9, 13,toupper(names(profiles)), pch =0, cex = 1.0, fill = 1:length(FeatImp_3))

```




